---
title: "Niagara report for cost of freight per shipment"
author: "Alberto Castillo"
date: "4/30/2018"
output: html_document
---


```{r include=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
options(scipen=999)
sku <- read.csv('sku.info.csv')
sales <- read.csv('Sales Q1 2017.csv') 
freight <- read.csv('Freight Q1 2017.csv')
setnames(freight, old=c('Destination.Customer','Sku','City','State','Zip.Code','Total.Spend', 'LOADTYPE', 'Shipment.24.Pk.Eq'), new=c('customer_name','sku','ship_city','ship_state','zip_code','total_spend', 'shipping_method', 'package_24_shipment'))                                                                
setnames(sku, old=c('Item.Number', 'X24.Equivalent', 'Cases.per.Truck'), new = c('sku', 'X24', 'cases_per_truck'))
setnames(sales, old = c('Party.Name', 'Segment1', 'Ship.City', 'Ship.State', 'Ship.Postal.Code', 'Shipping.Method', 'Cases.SUM', 'X24.PBE.Cases'), 
         new=c('customer_name','sku','ship_city','ship_state','zip_code','shipping_method','cases_sold', 'package_24_shipment'))
freight_cost <- select(freight,
                       customer_name,
                       sku,
                       ship_city,
                       ship_state,
                       zip_code,
                       shipping_method,
                       package_24_shipment,
                       total_spend)
freight_cost$ship_city <- as.character(freight_cost$ship_city)
freight_cost[freight_cost$ship_city == 'TLAJOMULCO DE ZUNIGA SAMS-6238', 'ship_city'] <- 'TLAJOMULCO DE ZUNIGA'
freight_cost[freight_cost$ship_city == 'CHALCO SAMS-7505', 'ship_city'] <- "CHALCO"
freight_cost[freight_cost$ship_city == 'CUAUTITLAN IZCALLI SAMS-6388', 'ship_city'] = "CUAUTITLAN IZCALLI"
freight_cost[freight_cost$ship_city == 'MONTERREY SAMS-4995', 'ship_city'] <- 'MONTERREY'
freight_cost[freight_cost$ship_city == 'CULIACAN SAMS-4971', 'ship_city'] <- 'CULIACAN'
freight_cost[freight_cost$ship_city == 'VILLAHERMOSA TAB. SAMS 6550', 'ship_city'] <- 'VILLAHERMOSA TABASCO'
freight_cost[freight_cost$ship_city == 'TLAJOMULCO DE ZUNIGA WM-7460', 'ship_city'] <- 'TLAJOMULCO DE ZUNIGA'
freight_cost[freight_cost$ship_city == 'TLAJOMULCO DE ZUNIGA WM-7493', 'ship_city'] <- 'TLAJOMULCO DE ZUNIGA'
freight_cost[freight_cost$ship_city == 'CHALCO WM-7459', 'ship_city'] <- 'CHALCO'
freight_cost[freight_cost$ship_city == 'CHALCO WM-7471', 'ship_city'] <- 'CHALCO'
freight_cost[freight_cost$ship_city == 'CUAUTITLAN IZCALLI WM-7457', 'ship_city'] <- 'CUAUTITLAN IZCALLI'
freight_cost[freight_cost$ship_city == 'CUAUTITLAN IZCALLI WM-7464', 'ship_city'] <- 'CUAUTITLAN IZCALLI'
freight_cost[freight_cost$ship_city == 'CUAUTITLAN IZCALLI WM-7482', 'ship_city'] <- 'CUAUTITLAN IZCALLI'
freight_cost[freight_cost$ship_city == 'CUAUTITLAN IZCALLI WM-7494', 'ship_city'] <- 'CUAUTITLAN IZCALLI'
freight_cost[freight_cost$ship_city == 'MONTERREY WM-7461', 'ship_city'] <- 'MONTERREY'
freight_cost[freight_cost$ship_city == 'MONTERREY WM-7490', 'ship_city'] <- 'MONTERREY'
freight_cost[freight_cost$ship_city == 'CULIACAN WM-7455', 'ship_city'] <- 'CULIACAN'
freight_cost[freight_cost$ship_city == 'CULIACAN WM-7487', 'ship_city'] <- 'CULIACAN'
freight_cost[freight_cost$ship_city == 'VILLAHERMOSA TAB. WM-7468', 'ship_city'] <- 'VILLAHERMOSA TABASCO'
freight_cost[freight_cost$ship_city == 'VILLAHERMOSA TABASCO WM 7453', 'ship_city'] <- 'VILLAHERMOSA TABASCO'
freight_cost <- filter(freight_cost, shipping_method != 'CustomerPickup')

sales_cost <- select(sales,
                     customer_name,
                     sku,
                     ship_city,
                     ship_state,
                     zip_code,
                     shipping_method,
                     package_24_shipment, 
                     cases_sold)
sales_cost$ship_city <- as.character(sales_cost$ship_city)
sales_cost$shipping_method <- as.character(sales_cost$shipping_method)
sales_cost[sales_cost$shipping_method == "DELIVERY", 'shipping_method'] <- 'Delivery'
sales_cost[sales_cost$shipping_method == 'Paid', 'shipping_method'] <- 'Delivery'
sales_cost[sales_cost$ship_city == "Ramos Arizpe", 'ship_city'] <- 'RAMOS ARIZPE'
sales_cost[sales_cost$ship_state == 'mx', 'ship_state'] <- 'MX'
sales_cost <- filter(sales_cost, shipping_method != 'PICKUP')
sales_cost$cases_sold <- abs(sales_cost$cases_sold)
freight_group <- freight_cost %>%
  group_by(customer_name, sku, ship_city, ship_state, zip_code, shipping_method) %>%
  summarise(package_24_shipment = sum(package_24_shipment), total_spend = sum(total_spend)) %>%
  ungroup()
sales_group <- sales_cost %>%
  group_by(customer_name, sku, ship_city, ship_state, zip_code, shipping_method) %>%
  summarise(package_24_shipment = sum(package_24_shipment), cases_sold = sum(cases_sold)) %>%
  ungroup()
sku_sold <- select(sales_group, sku, cases_sold) %>%
  group_by(sku) %>%
  summarise(cases_sold = sum(cases_sold)) %>%
  ungroup()
sku_freight <- select(freight_group, sku, total_spend) %>%
  group_by(sku) %>%
  summarise(total_spend = sum(total_spend)) %>%
  ungroup()
sku_total <- merge(sku_sold, sku_freight, by =c('sku')) %>%
  mutate(avg_spend = total_spend/cases_sold)
sku_total <- select(sku_total, sku, avg_spend)
sales_final <- merge(sales_group, sku_total, by =c('sku'))
sales_final$zip_code <- as.character(sales_final$zip_code)
sales_final$shipping_method <- as.factor(sales_final$shipping_method)
sales_final$ship_city <- as.factor(sales_final$ship_city)
sales_final$zip_code <- as.factor(sales_final$zip_code)
world.map <- map_data("world")
mexico <- subset(world.map, region == "Mexico")
select_stuff <- select(sales_final,
                       ship_state,
                       ship_city,
                       zip_code,
                       cases_sold) %>%
  group_by(ship_city, ship_state, zip_code) %>%
  summarise(avg_case = mean(cases_sold)) %>%
  ungroup()
select_stuff['long'] <- NA
select_stuff['lat'] <- NA
select_stuff$long <- as.double(c('-98.85', '-104.0023', '-100.1471', '-99.1783', '-99.1783', '-99.1972',
                                 '-107.4631', '-103.3167', '-100.3122', '-103.6167', '-100.2138', '-99.4389',
                                 '-92.5124', '-99.1833', '-103.3167', '-103.1833', '-99.4919', '-92.8138',
                                 '-93.4', '-92.8'))
select_stuff$lat <- as.double(c('19.25', '19.1116', '25.7636', '19.6997', '19.6997', '19.5936',
                                '24.7733', '20.5667', '25.8002', '25.6333', '25.7070', '24.9761',
                                '15.9266', '19.7456', '20.4333', '20.5167', '19.6651', '17.9278',
                                '17.8667', '17.5458'))
select_imp <- select(select_stuff,
                     ship_city,
                     zip_code,
                     long,
                     lat)
sales_final2 <- merge(sales_final, select_imp, by = c('ship_city', 'zip_code')) 
  
cities <- select(sales_final, ship_city, cases_sold) %>%
  group_by(ship_city) %>%
  summarise(total_sold = sum(cases_sold)) %>%
  ungroup()
proprtion <- function(x, removeNA=TRUE){
  nc <- length(x) #length of the column
  prop <- numeric(nc) #We need to initialize this vector. We know this will be numeric so we add that now.
  for (i in 1:nc){
    prop[i] <- x[i] / sum(x)
  }
  prop
}
prop <- select(cities, ship_city, total_sold)
prop$prop <- NULL
prop$prop <- proprtion(cities$total_sold)
prop$prop <- formatC(prop$prop, digits = 5, format = "f")
prop$prop <- as.double(prop$prop)

library(plotly) #for pie chart
pie <- plot_ly(prop, labels = prop$ship_city, values = prop$prop, type = 'pie',textposition = 'outside',textinfo = 'label+percent') %>%
  layout(title = 'Percentage of Cases Sold to each Shipping City',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)) 



```
## Summary of out Data
Here is a quick summary of our data. We see how many times values occured in our data frame as well as what were the quartiles values for our data.
```{r}
summary(sales_final)
```

## Visulization of Customers
Now lets take a much closer look and see in which cities customers are buying these water cases.
***
```{r fig.width=10}
ggplot(sales_final, aes(x = customer_name, fill = ship_city))+
  geom_bar(color = 'black')+
  labs(x = 'Customer',  y= 'Where that customer is buying from', fill = 'City')

```

## Visualization of Cases Sold
Here we show which city had the most sold cases. What was the breakdown of the shipments to each city.
We can see which cities are the ones who are buying the most cases and which ones are buying the average amount of around 47306 cases. We can make cost efficient decisions in the next quarter so to maximize our orders with those cities that are buying the most cases.
```{r fig.width= 10}
#This is the total amount of cases that were shipped to each city
print(cities)

#Lets show the weight of total cases sold to each city
pie

ggplot(sales_final, aes(x = cases_sold, fill = ship_city))+
  geom_histogram(color = 'black')+
  labs(x = 'Cases Sold', fill = "City")

```

## Showing our Data on a Map
Here we see that I divided up the information and plotted it on a map of Mexico the busiest place that the cases were shipped for Q1 of 2017. This is good to visualize where we can focus more of the shipments in the future as to be sure to maximize shipments and profits. 
```{r fig.width= 10, fig.height= 8}
ggplot(mexico, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill = "white",
               color = "gray40", size = .2) +
  geom_point(data = sales_final2, aes(x = long, y = lat, size = cases_sold, color = ship_city))+
  scale_size_continuous(range = c(8,15))+
  labs(color = 'City of Destination', size = "Average cost of shipment", title ='Comparing the volume of shipment to each city in Mexico')

```

## Lets look at the Distribution of Shipment Cost to each City
Here we have a boxplot that shows us the median price of shipping across all the cities in Q1. We see what the maximum value and minimum value were for shipping to that respective city. The average cost of shipment was $10.95.
```{r fig.width=5}
ggplot(sales_final, aes(x = reorder(ship_city, avg_spend, median), y = avg_spend))+
  geom_boxplot()+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(x = "City", y = "Shipment Cost per Case(in Dollars)")
```


## Regression between Cost per Case to Chip, The Amount of Cases Shipped, and 24 Pack Cases
Here we are showing the average shipping price of shipping cases of water against how many cases where sold at that price point. We then want to see how many complete 24 packs of water of those cases sold compare against to the price of the shipment. Since different products have different bottle sizes and pack sizes, so “24-pack equivalent” cases are used as a scale for a comparison between different product types.
```{r}
ggplot(sales_final, aes(x = avg_spend, y = cases_sold))+
  geom_point(aes(size = package_24_shipment), alpha = 1/3)+
  labs(x = 'freight cost', y = 'cases sold', size = '24-pack equivalents')
```


Now we want to smooth out this data to see what sort of relationships each state state has with regards to the cases sold, shipping cost, and 24 pack equivalents. We can determine what relationship our data has form this information. As cases start to get larger we see which states comsume the most of our product, while others start to fall off or comsume less. We can also see at what amount states start to increase their consumptions. This informations is valuable as it would lead to know how many cases to estimate to each state which can result in having a more efficient second quarter.
```{r}
ggplot(sales_final, aes(x = avg_spend, y = cases_sold))+
  geom_point(aes(size = package_24_shipment), alpha = 1/3)+
  geom_smooth(aes(color = ship_state), se = FALSE) +
  labs(x = 'freight cost', y = 'cases sold', size = '24-pack equivalents')
```


## Conclusion

With this data we can now make accurate estimations for quarter 2 of sales and freight. We can estimate how much it would cost to ship to these same locations in quarter 2 as well as know about how much each respective state and city will consume in the future. The data also tells us what the perfect amount of cases would be for each state as to now have them start to consume less product; therefore, making our shipment process much more efficient as we will know how much a specific state/city will expect to consume.
